\textbf{ Accurate spectroscopic estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-24 paper/sp.tex

w%% This file is part of the spectroscopic_parallax project
%% Copyright 2018 the authors.

@@

Since the photometry is (by the ``good features'' assumption above)
better measured than the parallaxes, we
presume that the parallax uncertainties dominate
the uncertainties in the luminosity measurements.
For each star $n$ in our training set, we define an inverse-square-root
luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
\begin{eqnarray}
Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
\quad,
\end{eqnarray}
where
$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
for star $n$.
In what follows, the $Q_n$ are the training data.
Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
\begin{eqnarray}
M_{Kn} &\approx& \frac{1}{0.2\,\ln 10}\,\ln Q_n
\quad.
\end{eqnarray}
These would be the absolute magnitudes of the stars under the
``the distance is the inverse of the
measured parallax'' assumption, although we never, in this work, ever make that
assumption, by God.
We construct the $Q_n$ because they are related directly to the absolute magnitudes,
but they (under assumptions) are subject to known Gaussian noise.

The log-likelihood function can be expressed heuristically as
\begin{eqnarray}
Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
\\
\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
\quad ,
\end{eqnarray}
where
the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
of some $D$-dimensional feature vector $x_n$,
$\theta$ is a $D$-vector of linear coefficients,
and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters
under the assumption of known Gaussian noise and
that there are $N$ independently measured stars $n$.
This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.

In addition, we assume that many entries in the $D$-vector will be zero
(the ``sparsity'' assumption).
In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
\begin{eqnarray}
\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi(\theta)^2 + ||\lambda\cdot\theta||_1^1\right]
\quad ,
\end{eqnarray}
where
$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
different parts of the $x_n$ vectors differently),
and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
This kind of regularization adds a convex term to the optimization and leads to
sparse optima.
We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
very coarse cross-validation.

@@

>>>>

%% This file is part of the spectroscopic_parallax project
%% Copyright 2018 the authors.

@@

Since the photometry is (by the ``good features'' assumption above)
better measured than the parallaxes, we
presume that the parallax uncertainties dominate
the uncertainties in the luminosity measurements.
For each star $n$ in our training set, we define an inverse-square-root
luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
\begin{eqnarray}
Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
\quad,
\end{eqnarray}
where
$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
for star $n$.
In what follows, the $Q_n$ are the training data.
Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
\begin{eqnarray}
M_{Kn} &\approx& 5\,\log_{10} Q_n
\quad.
\end{eqnarray}
These would be the absolute magnitudes of the stars under the
``the distance is the inverse of the
measured parallax'' assumption, although we never, in this work, ever make that
assumption, by God.
We construct the $Q_n$ because they are related directly to the absolute magnitudes,
but they (under assumptions) are subject to known Gaussian noise.

The log-likelihood function can be expressed heuristically as
\begin{eqnarray}
Q_n &=& \exp(\theta\cdot x_n) + \mbox{noise}
\\
\chi^2(\theta) &\equiv& \sum_{n=1}^N \frac{[Q_n - \exp(\theta\cdot x_n)]^2}{\sigma_n^2}
\quad ,
\end{eqnarray}
where
the model is that the $\ln Q_n$ can be expressed as a linear combination of the components
of some $D$-dimensional feature vector $x_n$,
$\theta$ is a $D$-vector of linear coefficients,
and $\chi^2(\theta)$ is the negative-log-likelihood for the parameters $\theta$
under the assumption of known Gaussian noise and
that there are $N$ independently measured stars $n$.
This feature vector $x_n$ contains photometry in a few bands, and all 8000-ish pixels
in the pseudo-continuum-normalized \apogee\ spectrum, so $D$ is on the order of 8000.

In addition, we assume that many entries in the $D$-vector will be zero
(the ``sparsity'' assumption).
In order to encourage this, we optimize not $\chi^2$ but a regularized objective function
\begin{eqnarray}
\hat{\theta} &\leftarrow& \argmin_{\theta}\left[\frac{1}{2}\,\chi^2(\theta) + ||\lambda\cdot\theta||_1^1\right]
\quad ,
\end{eqnarray}
where
$\lambda$ is a $D$-vector of regularization parameters (such that we can regularize
different parts of the $x_n$ vectors differently),
and $||x||_1^1$ is the L1-norm or sum of absolute values of the components of $x$.
This kind of regularization adds a convex term to the optimization and leads to
sparse optima.
We will set the value $\lambda$ of the $D$-vector of regularization strengths by a
very coarse cross-validation.

Once the model is optimized, the prediction for a new stellar absolute magnitude
$M_{Km}$ for a star $m$ in the validation or test set is
\begin{eqnarray}
Q_m &=& \exp(\theta\cdot x_m)
\\
M_{Km} &=& 5\,\log_{10} Q_m
\quad ,
\end{eqnarray}
where $x_m$ is the feature $D$-vector for star $m$.
Thus the model can be trained on a training set of stars and applied to make
predictions for a validation set or a test set.

@@

What did we do and what's cool?

Why did we make the assumptions we made and what did they gain us and cost us?

What is this useful for and how do you use them safely?

(Mention independence of every result from the \gaia\ astrometry.)

Come back to SDSS-V and show the rotation curve of the Galaxy.
